---
title: "p8105_hw2_KAB2310"
author: "Kamiah Brown"
date: "2024-10-01"
output: github_document
---
## This is my submission for Homework 2.

#Set Up 
```{r}
library(tidyverse)
library(dplyr)
library(readxl)
```
#Problem 1
```{r}
nyc_transit_df = 
  read_csv("NYC_Transit_Subway_Entrance_And_Exit_Data.csv", na = c("NA", "", "."))
```
#Cleaning data
```{r}
nyc_transit_clean = nyc_transit_df |>
  janitor::clean_names() |>
  select(
    line = line, 
    station = station_name, 
    name = station_name, 
    station_latitude = station_latitude, 
    station_longitude = station_longitude, 
    route1, route2, route3, route4, route5, route6, route7, route8, route9, route10, route11,
    entry = entry, 
    vending = vending, 
    entrance_type = entrance_type, 
    ada = ada
  ) |>
  
  unite(routes_served, route1:route11, sep = ", ", na.rm = TRUE) |>
  mutate(entry = ifelse(entry == "yes", TRUE, FALSE))

nrow(nyc_transit_clean)
ncol(nyc_transit_clean)

```
The nyc transit clean data set has 1868 observations and 10 variables. 
The data set contain the following variables: line, station, name, station latitude, station longitude, routes served, entry, vending, entrance type and ada. I selected and retained the 10 variables. Additionally, I combined the 11 routes in using the unite() function to create a new column called "route_served". The size of data set is `r nrow(nyc_transit_clean)`  and `r ncol(nyc_transit_clean)`.
Since each row in the data set corresponds with the correct attribute, I believe the data is tidy. There are no missing variables from previewing the nyc transit clean data set. 


#Answer the following questions using these data:

How many distinct stations are there? Note that stations are identified both by name and by line (e.g. 125th St 8th Avenue; 125st Broadway; 125st Lenox); the distinct function may be useful here.

```{r}
distinct_stations = nyc_transit_clean |>
  distinct(station, line)

distinct_stations =
  nrow(distinct_stations)

distinct_stations
```
There are 465 distinct stations. 

How many stations are ADA compliant?
```{r}
ada_compliant_stations =  nyc_transit_clean |>
  filter(ada == TRUE) |>
  distinct(station, line) |> 
  nrow()

ada_compliant_stations
```
There are 84 stations with ADA compliant.

What proportion of station entrances / exits without vending allow entrance?
```{r}
without_vending_data = nyc_transit_df |>
  filter(Vending == "NO")

allow_entry_proportion <- without_vending_data %>%
  summarise(proportion = mean(Entry == "YES")) %>%
  pull(proportion)

allow_entry_proportion

```
37.7% of station entrances or exits that do not have vending machines still allow people to enter the station.

Reformat data so that route number and route name are distinct variables.
```{r}
nyc_transit_reformat = nyc_transit_clean |>
  separate_rows(routes_served, sep = ", ")  
view (nyc_transit_reformat)

a_train_stations = nyc_transit_reformat |>
  filter(routes_served == "A") |>
  distinct(station, line)  

 nrow(a_train_stations)
 
 ada_compliant_a_train_stations = nyc_transit_reformat |>
  filter(routes_served == "A", ada == TRUE) |>
  distinct(station, line)
 
 nrow(ada_compliant_a_train_stations)
```
There are `r nrow(a_train_stations)` distinct stations that serve the A train. Of the 60, `r nrow(ada_compliant_a_train_stations)` are ADA complaint.

#Problem 2
```{r}
# Mr Trash Wheel sheet
mr_trash_wheel = read_excel("202309 Trash Wheel Collection Data.xlsx", 
                             sheet = "Mr. Trash Wheel", 
                             skip = 1) |>
  janitor::clean_names() |>
  filter(!is.na(dumpster)) |>
  mutate(
    sports_balls = as.integer(round(sports_balls)), 
    year = as.character(year),  
    trash_wheel = "Mr. Trash Wheel"
  )
mr_trash_wheel = mr_trash_wheel |>
  select(where(~ !all(is.na(.))))

view(mr_trash_wheel)

#There are 584 observations and 15 variables in Mr. Trash Wheel Data set. Some of the key variables consist of dumpster, month, year, date, weight tons, volume cubic yards, plastic bottles, polystyrene, cigarette butts, glass bottles, plastic bottles, wrappers, sports balls and homes powered. Extra columns (2)  were accidentally added in the data manipulation process, so I removed them in lines 122-123. 


#Professor Trash Wheel sheet
prof_trash_wheel = read_excel("202309 Trash Wheel Collection Data.xlsx", 
                               sheet = "Professor Trash Wheel", 
                               skip = 1) |>
  janitor::clean_names() |>
  filter(!is.na(dumpster)) |>
  mutate(
     year = as.character(year),  
    trash_wheel = "Professor Trash Wheel"
  )

#There are 106 observations and 14 variables. Some of the key variables consist of dumpster, month, year, date, weight tons, volume cubic yards, plastic bottles, polystyrene, cigarette butts, glass bottles, plastic bags, wrappers, home powered, and trash wheel. 

#Gwynnda Trash Wheel sheet
gwynnda_trash_wheel = read_excel("202309 Trash Wheel Collection Data.xlsx", 
                                  sheet = "Gwynnda Trash Wheel", 
                                  skip = 1) |>
  janitor::clean_names() |>
  filter(!is.na(dumpster)) |>
  mutate(
    year = as.character(year),  
    trash_wheel = "Gwynnda Trash Wheel"
  )

#There are 155 observations and 13 variables. Some of the key variables consist of dumpster, month, year, date, weight tons, volume cubic yards, plastic bottles, polystyrene, cigarette butts, plastic bags, wrappers, homes powered, and trash wheel.

# Combine sheets
trash_wheels_df = 
  bind_rows(
  mr_trash_wheel,
  prof_trash_wheel,
  gwynnda_trash_wheel
)

trash_wheels_df = trash_wheels_df |>
  relocate(trash_wheel, sports_balls, .before = everything()) 

view(trash_wheels_df)
```


```{r}
#Finding total weight of trash collected by Prof Trash Wheel; I originally came across an error stated "character string is not in a standard unambiguous format", so I converted the date column into the correct formatting. 
trash_wheels_df =
  trash_wheels_df |>
  mutate(date = as.Date(date, format = "%Y-%m-%d")) 

prof_trash_data = 
  trash_wheels_df |>
  filter(trash_wheel == "Professor Trash Wheel")

total_weight_prof_trash =
  prof_trash_data |>
  summarize(total_weight = sum(weight_tons, na.rm = TRUE))

total_weight_prof_trash

#Finding the total number of cig butts collected by Gwynnda in June of 2022. 
gwynnda_june2022 = trash_wheels_df |>
  filter(trash_wheel == "Gwynnda Trash Wheel", 
         year == 2022, 
         month == 6)

total_cig_butts = gwynnda_june2022 |>
  summarize(total_cigarette_butts = sum(cigarette_butts, na.rm = TRUE))

total_cig_butts
```

The size of the resulting data set, trash_wheels_df is  `r nrow(trash_wheels_df)` and `r ncol(trash_wheels_df). Some of the key variables are trash wheel, sports balls, dumpster, month, year, date, weight tons, volume cubic yards, plastic bottles, polystyrene, cigarette butts, glass bottles, plastic bottles, wrappers and homes powered. 
The total weight of trash collected by Professor Trash Wheel is 216.
The total number of cigarette butts collected by Gwynnda in June of 2022 is 0. 

#Problem 3
```{r}
bakers = read_csv("bakers.csv")
bakes = read_csv("bakes.csv")
results = read_csv("results.csv")

bakers =
  bakers |> rename(baker = 'Baker Name')
bakes = 
  bakes |> rename(baker = 'Baker')

results = results |>
  select(series, episode, baker, technical, result)

results = results |> 
  rename(series = `...1`, episode = `...2`, baker = `...3`, technical = `...4`, result = `IN = stayed in; OUT = Eliminated; STAR BAKER = Star Baker; WINNER = Series Winner; Runner-up = Series Runner up; WD = withdrew`)

results = results |>
  select(series, episode, baker, technical, result) #Error in select

# Merge datasets
bakeoff_df <- bakes |>
  left_join(bakers, by = "baker") |>
  left_join(results, by = c("baker", "series"))


